---
title: "R Notebook"
output: html_notebook
params:
  serie: '5ef'
---


```{r}
library(tidyverse)
library(caret)
library(rpart)
library(rpart.plot)

```

# Import 
```{r modelo v1}
saeb <- read.csv2(paste0("../output/books/base",params$serie,".csv"))

df <-
  saeb %>% 
  filter(id_dependencia_adm == 'Municipal') %>% 
    select(
      id_escola,
      media,
      nivel_socio_economico,
      starts_with('tx'),
      ends_with('ipvs'),
      ends_with('imp')
    ) %>% 
    mutate(
      nivel_socio_economico = fct_collapse(nivel_socio_economico,
                                                IV = c("Nível II","Nível III","Nível IV"),
                                                V = c("Nível V",""),
                                                VI = c("Nível VI","Nível VII")
                                                ))
```

# Pilares
## Pilar Família
```{r}
rpart_familia <- rpart(media ~ tx_resp_q003a_B + tx_resp_q003b_B + tx_resp_q003c_B + tx_resp_q003d_B + tx_resp_q003e_B,
                       data = df)
rpart.plot(rpart_familia)
```

### Cria Pilar
```{r}
df <-  df %>% 
  mutate(pred_familia = round(predict(rpart_familia, df)),
         pilar_familia = case_when(
           pred_familia >= 243 ~ 'A',
           pred_familia >= 236 ~ 'B',
           pred_familia >= 231 ~ 'C',
           pred_familia >= 228 ~ 'D',
           pred_familia == 220 ~ 'E',
         ))

table(df$pilar_familia)
```
### R2
```{r}
lm_familia <- lm(media ~ tx_resp_q003a_B + tx_resp_q003b_B + tx_resp_q003c_B, df)
summary(lm_familia)$r.squared
```

```{r}
lm_familia <- lm(media ~ pilar_familia, df)
summary(lm_familia)$r.squared
```

## Pilar Letramento
### Rpart
```{r}
rpart_letramento <- rpart(media ~ .,
                       data = df %>% select(media,starts_with("tx_resp_q004"),starts_with("tx_resp_q005"))
                       ,minbucket = 0.05*nrow(df)
                       )
rpart.plot(rpart_letramento)
```

### Cria Pilar
```{r}
df <-  df %>% 
  mutate(pred_letramento = round(predict(rpart_letramento, df)),
         pilar_letramento = case_when(
           pred_letramento >= 249 ~ 'A',
           pred_letramento >= 237 ~ 'B',
           pred_letramento >= 232 ~ 'C',
           pred_letramento >= 229 ~ 'D',
           pred_letramento >= 223 ~ 'E',
         ))

table(df$pilar_letramento)
```

### R2
```{r}
model <- lm(media ~ tx_resp_q004_A +  tx_resp_q004_E + tx_resp_q005_E, df)
summary(model)$r.squared
```

```{r}
model <- lm(media ~ pilar_letramento, df)
summary(model)$r.squared
```



## Pilar Região
### Rpart
```{r}
rpart_regiao <- rpart(media ~ tx_resp_q008a_B + tx_resp_q008b_B + tx_resp_q008c_B,
                       data = df
                       ,minbucket = 0.05*nrow(df)
                       )
rpart.plot(rpart_regiao)
```

### Cria Pilar
```{r}
df <-  df %>% 
  mutate(pred_regiao = round(predict(rpart_regiao, df)),
         pilar_regiao = case_when(
           pred_regiao == 244 ~ 'A',
           pred_regiao == 236 ~ 'B',
           pred_regiao == 229 ~ 'C',
           pred_regiao == 223 ~ 'D'
         ))

table(df$pilar_regiao)
```

### R2
```{r}
model <- lm(media ~ tx_resp_q008a_B + tx_resp_q008c_B, df)
summary(model)$r.squared
```

```{r}
model <- lm(media ~ pilar_regiao, df)
summary(model)$r.squared
```




## Pilar Localização
### Rpart
```{r}
rpart_localizacao <- rpart(media ~ .,
                       data = df %>% select(media,starts_with('tx_resp_q011'),starts_with('tx_resp_q012'))
                       ,minbucket = 0.05*nrow(df)
                       )
rpart.plot(rpart_localizacao)
```

### Cria Pilar
```{r}
df <-  df %>% 
  mutate(pred_localizacao = round(predict(rpart_localizacao, df)),
         pilar_localizacao = case_when(
           pred_localizacao == 251 ~ 'A',
           pred_localizacao == 240 ~ 'B',
           pred_localizacao == 232 ~ 'C',
           pred_localizacao == 230 ~ 'D',
           pred_localizacao == 225 ~ 'E',
           pred_localizacao == 219 ~ 'F'
         ))

table(df$pilar_localizacao)
```

### R2
```{r}
model <- lm(media ~ tx_resp_q012_D + tx_resp_q012_F, df)
summary(model)$r.squared
```

```{r}
model <- lm(media ~ pilar_localizacao, df)
summary(model)$r.squared
```




## Pilar Bens 
### Rpart
```{r}
rpart_bens <- rpart(media ~ .,
                       data = df %>% select(media,starts_with('tx_resp_q009'),starts_with('tx_resp_q010'))
                       ,minbucket = 0.05*nrow(df)
                       )
rpart.plot(rpart_bens)
```

### Cria Pilar
```{r}
df <-  df %>% 
  mutate(pred_bens = round(predict(rpart_bens, df)),
         pilar_bens = case_when(
           pred_bens == 252 ~ 'A',
           pred_bens == 242 ~ 'B',
           pred_bens == 236 ~ 'C',
           pred_bens == 233 ~ 'D',
           pred_bens == 225 ~ 'E',
           pred_bens == 223 ~ 'F',
           pred_bens == 215 ~ 'G',
         ))

table(df$pilar_bens)
```

### R2
```{r}
model <- lm(media ~ pilar_bens, df)
summary(model)$r.squared
```




## Pilar Bens Supérfluos
### Rpart
```{r}
rpart_bens <- rpart(media ~ .,
                       data = df %>% select(media,
                                            starts_with('tx_resp_q009b'),
                                            starts_with('tx_resp_q009c'),
                                            starts_with('tx_resp_q009e'),
                                            starts_with('tx_resp_q009g'))
                       ,minbucket = 0.05*nrow(df)
                       )
rpart.plot(rpart_bens)
```

### Cria Pilar
```{r}
df <-  df %>% 
  mutate(pred_bens = round(predict(rpart_bens, df)),
         pilar_bens_superfluos = case_when(
           pred_bens >= 248 ~ 'A',
           pred_bens >= 240 ~ 'B',
           pred_bens >= 237 ~ 'C',
           pred_bens >= 230 ~ 'D',
           pred_bens >= 228 ~ 'E',
           pred_bens >= 225 ~ 'F',
           pred_bens >= 217 ~ 'G'
         ))

table(df$pilar_bens_superfluos)
```

### R2
```{r}
model <- lm(media ~ pilar_bens_superfluos, df)
summary(model)$r.squared
```




## Pilar Bens serviços
### Rpart
```{r}
rpart_bens <- rpart(media ~ .,
                       data = df %>% select(media,
                                            starts_with('tx_resp_q007'),
                                            starts_with('tx_resp_q010a'),
                                            starts_with('tx_resp_q010b'))
                       ,minbucket = 0.05*nrow(df)
                       )
rpart.plot(rpart_bens)
```

### Cria Pilar
```{r}
df <-  df %>% 
  mutate(pred_bens = round(predict(rpart_bens, df)),
         pilar_bens_servicos = case_when(
           pred_bens >= 245 ~ 'A',
           pred_bens >= 237 ~ 'B',
           pred_bens >= 234 ~ 'C',
           pred_bens >= 228 ~ 'D',
           pred_bens >= 222 ~ 'E',
           pred_bens >= 214 ~ 'F',
         ))

table(df$pilar_bens_servicos)
```

### R2
```{r}
model <- lm(media ~ pilar_bens_servicos, df)
summary(model)$r.squared
```





## Pilar Bens Cômodos
### Rpart
```{r}
rpart_bens <- rpart(media ~ .,
                       data = df %>% select(media,
                                            starts_with('tx_resp_q009d'),
                                            starts_with('tx_resp_q009f'),
                                            starts_with('tx_resp_q010c'),
                                            starts_with('tx_resp_q010d'),
                                            starts_with('tx_resp_q010e'))
                       ,minbucket = 0.05*nrow(df)
                       )
rpart.plot(rpart_bens)
```

### Cria Pilar
```{r}
df <-  df %>% 
  mutate(pred_bens = round(predict(rpart_bens, df)),
         pilar_bens_comodos = case_when(
           pred_bens >= 251 ~ 'A',
           pred_bens >= 242 ~ 'B',
           pred_bens >= 237 ~ 'C',
           pred_bens >= 232 ~ 'D',
           pred_bens >= 225 ~ 'E',
           pred_bens >= 224 ~ 'F',
           pred_bens >= 217 ~ 'G',
         ))

table(df$pilar_bens_comodos)
```

### R2
```{r}
model <- lm(media ~ pilar_bens_comodos, df)
summary(model)$r.squared
```





## Pilar Bens Eletrodomésticos
### Rpart
```{r}
rpart_bens <- rpart(media ~ .,
                       data = df %>% select(media,
                                            starts_with('tx_resp_q009a'),
                                            starts_with('tx_resp_q010f'),
                                            starts_with('tx_resp_q010g'),
                                            starts_with('tx_resp_q010h'),
                                            starts_with('tx_resp_q010i'))
                       ,minbucket = 0.05*nrow(df)
                       )
rpart.plot(rpart_bens)
```

### Cria Pilar
```{r}
df <-  df %>% 
  mutate(pred_bens = round(predict(rpart_bens, df)),
         pilar_bens_eletrodomesticos = case_when(
           pred_bens >= 252 ~ 'A',
           pred_bens >= 242 ~ 'B',
           pred_bens >= 234 ~ 'C',
           pred_bens >= 233 ~ 'D',
           pred_bens >= 225 ~ 'E',
           pred_bens >= 216 ~ 'F'
         ))

table(df$pilar_bens_eletrodomesticos)
```

### R2
```{r}
model <- lm(media ~ pilar_bens_eletrodomesticos, df)
summary(model)$r.squared
```

# Pilar IPVS
### Rpart
```{r}
rpart_ipvs <- rpart(media ~ .,
                       data = df %>% select(media,
                                            ends_with('ipvs'))
                       ,minbucket = 0.05*nrow(df)
                       )
rpart.plot(rpart_ipvs)
```

```{r}
df <-  df %>% 
  mutate(pred_ipvs = round(predict(rpart_ipvs, df)),
         pilar_ipvs = case_when(
           pred_ipvs >= 247 ~ 'A',
           pred_ipvs >= 243 ~ 'B',
           pred_ipvs >= 241 ~ 'C',
           pred_ipvs >= 235 ~ 'D',
           pred_ipvs >= 230 ~ 'E',
           pred_ipvs >= 226 ~ 'F',
           pred_ipvs >= 217 ~ 'G'
         ))

table(df$pilar_ipvs)
```
### R2
```{r}
model <- lm(media ~ pilar_ipvs, df)
summary(model)$r.squared
```

# Pilar Seade
### Rpart
```{r}
rpart_seade <- rpart(media ~ .,
                       data = df %>% select(media,
                                            ends_with('imp'))
                       ,minbucket = 0.05*nrow(df)
                       )
rpart.plot(rpart_seade)
```
```{r}
df <-  df %>% 
  mutate(pred_seade = round(predict(rpart_seade, df)),
         pilar_seade = case_when(
           pred_seade >= 239 ~ 'A',
           pred_seade >= 236 ~ 'B',
           pred_seade >= 233 ~ 'C',
           pred_seade >= 224 ~ 'D',
           pred_seade >= 220 ~ 'E'
           
         ))

table(df$pilar_seade)
```
### R2
```{r}
model <- lm(media ~ pilar_seade, df)
summary(model)$r.squared
```


# Export
```{r}
pilar <- df %>% 
  select(id_escola,
         starts_with('pilar'))

head(pilar)
```

```{r}
url <- paste0("../output/pilares/pilar_municipais_",params$serie,".csv")
write.csv2(pilar, url, row.names = F, na = "")
```


