---
title: "R Notebook"
output: html_notebook
params:
  serie: '3em'
---


```{r}
library(tidyverse)
library(caret)
```

# Modelo v1: 0,701
```{r modelo v1}
df_publico <- read.csv2("output/books/df_publico.csv", na = "")

df <- df_publico %>% 
  filter(id_dependencia_adm == 'Estadual',
         id_serie == toupper(params$serie)) %>% 
  select(id_escola, 
         co_distrito,
         id_municipio,
         tipo_publico,
         media,
         nivel_socio_economico)
head(df)
```
## Nivel_socioeconomico
```{r var1}
df <- df %>% 
  mutate(nivel_socio_economico = as.character(nivel_socio_economico),
        nivel_socio_economico = 
               factor(case_when(
                 nivel_socio_economico %in% c('Nível II','Nível III') ~ 'Nível IV',
                 nivel_socio_economico %in% c('Nível VII') ~ 'Nível VI',
                         TRUE ~ nivel_socio_economico)))
```



# INSE Saresp: 0,706 e 0.702(input by group)
```{r eval=FALSE}
inse <- read_csv2("data/saresp/INSE_Geral 2018_1_0.csv")
colnames(inse) <- c("codesc","nome","dir","mun","inse")
inse <- inse %>% 
  transmute(id_escola = as.integer(paste0("35",str_pad(codesc,6,pad = "0"))),
            inse)
head(inse)
```

```{r eval=FALSE}
teste <- df %>% 
  left_join(inse, by = "id_escola")
summary(teste$inse)
teste %>% group_by(nivel_socio_economico) %>% summarise(inse = mean(inse, na.rm = T))
```

```{r eval=FALSE}
teste <- teste %>% 
  #mutate(inse = replace_na(inse, mean(inse, na.rm = T)))
  group_by(nivel_socio_economico) %>%
    mutate(
        inse = replace_na(inse, mean(inse, na.rm = T))
    ) %>% 
  ungroup()
summary(teste$inse)
teste %>% group_by(nivel_socio_economico) %>% summarise(inse = mean(inse, na.rm = T))
```

```{r eval=FALSE}
set.seed(231192)
model_ranger <- train(media ~ nivel_socio_economico + tx_resp_q017e_A + tx_resp_q005_F * tx_resp_q004_F * tx_resp_q004_E + tx_resp_q002_A + tx_resp_q010d_B + tx_resp_q009c_A + tx_resp_q012_A + tx_resp_q010i_B +              nascidos_vivos_mae18_2019_imp + inse,
                        data = teste %>% select(-id_escola),
                        method = "lm",
                        metric = "Rsquared",
                        trControl = trControl)


results$rf <- model_ranger$resample$Rsquared
results
mean(results$rf)
```

## IMport
## Pilares
```{r}
pilar <- read_csv2(paste0("output/pilares/pilar_", params$serie,".csv"))
head(pilar)
```
## Saresp: 0.6999666
```{r import saresp prop}
saresp <- read_csv2(paste0("output/books/saresp", params$serie,"_prop.csv"))
head(saresp)
```
## Saeb Book
```{r Import saeb}
url <- paste0("output/books/saeb", params$serie, "_prop.csv")
saeb <- read.csv2(url, na = "")
head(saeb)
```

## Seade
```{r}
seade <- read_csv2(paste0("output/books/seade", params$serie,".csv"))
```

## IPVS
```{r}
ipvs <- read_csv2(paste0("output/books/ipvs", params$serie,".csv"))
```

# Integra
```{r}
df <- df %>% 
  left_join(pilar, by = "id_escola") %>% 
  left_join(saeb, by = "id_escola") %>% 
  left_join(saresp, by = "id_escola") %>% 
  left_join(ipvs, by = "co_distrito") %>% 
  left_join(seade, by = "id_municipio") %>% 
  select(-co_distrito, -id_municipio)
dim(df)
```

# Checa Missing
```{r pilar}
df %>% select(starts_with('pilar')) %>% anyNA()
```
```{r saeb}
df %>% select(starts_with('tx')) %>% anyNA()
```
```{r saresp}
df %>% select(ends_with('spp')) %>% anyNA()
```

```{r ipvs}
df %>% select(ends_with('ipvs')) %>% anyNA()
```

```{r seade}
df %>% select(ends_with('imp')) %>% anyNA()
```
# Input Missing
```{r}
df2 <- df %>% 
  mutate(across(ends_with('spp'),~ifelse(is.na(.x), mean(.x, na.rm = TRUE), .x)))

df2 %>% select(ends_with('spp')) %>% anyNA()
```
```{r}
df3 <- df2 %>% 
  mutate(across(ends_with('imp'),~ifelse(is.na(.x), mean(.x, na.rm = TRUE), .x)))

df3 %>% select(ends_with('imp')) %>% anyNA()
```

# Modelos
## CV
```{r createfolds}
df <- df3 %>% filter(tipo_publico != 'ETEC') %>% 
  select(media,
         nivel_socio_economico,
         starts_with('pilar'),
         starts_with('tx'),
         ends_with('spp'),
         ends_with('ipvs'),
         ends_with('imp'))

set.seed(231192)
myFolds <- createFolds(df$media, k = 5)
trControl <- trainControl(
  method = "cv",
  number = 5,
  index = myFolds
)

results <- data.frame(fold = c('Fold 1', 'Fold 2', 'Fold 3', 'Fold 4', 'Fold 5'))

```

## v1
1. Com ETEC: 0.7010868
2. Sem ETEC: 0.5700405
```{r}
set.seed(231192)
model_v1 <- train(media ~nivel_socio_economico + tx_resp_q017e_A + tx_resp_q005_F * tx_resp_q004_F * tx_resp_q004_E + tx_resp_q002_A + tx_resp_q010d_B + tx_resp_q009c_A + tx_resp_q012_A + tx_resp_q010i_B +              nascidos_vivos_mae18_2019_imp,
                        data = df3,
                        method = "lm",
                        metric = "Rsquared",
                        trControl = trControl)


model_v1$resample$Rsquared
mean(model_v1$resample$Rsquared)
```

## v2
1. nivel_socio_economico + pilar_familia + pilar_letramento + pilar_regiao + pilar_localizacao + pilar_bens: 0.5287162
2. tx_resp_q017e_A: 0.5524336
3. tx_resp_q002_A: 0.5712744
0. pilar_ipvs: 0.5690674
0. ipvs567_ipvs: 0.5697283
4. pilar_saresp: 0.577281
0. pilar_seade: 0.5746202
5. ind_envelhecimento_2019_imp:  0.5782496
0. trocar pilar bens por superfluos e comodos: 0.5713985
6. tx_resp_q017c_A

### testes com pilar: 0.581911
1. familia:
tx_resp_q003a_B+tx_resp_q003b_B+tx_resp_q003c_B: 0.5812321
2. letramento 
tx_resp_q004_E+tx_resp_q005_E + tx_resp_q005_F: 0.5917858
3. regiao
tx_resp_q008b_B: 0.5709533
4. localizacao 
tx_resp_q011_B + tx_resp_q012_A+ tx_resp_q012_F: 0.5822252 
5. bens
tx_resp_q009c_A+tx_resp_q010c_B+tx_resp_q010d_B+tx_resp_q010i_B: 0.5839408
```{r}
set.seed(231192)
model_v2 <- train(media ~ nivel_socio_economico + pilar_familia + tx_resp_q004_E+tx_resp_q005_E + tx_resp_q004_F + tx_resp_q005_F + pilar_regiao + pilar_localizacao + pilar_bens + tx_resp_q017c_A +
  + I(tx_resp_q017e_A^2) + tx_resp_q002_A + pilar_saresp,
                        data = df3,
                        method = "lm",
                        metric = "Rsquared",
                        trControl = trControl)


model_v2$resample$Rsquared
mean(model_v2$resample$Rsquared)
```

# Teste com Random Forest
```{r}
set.seed(231192)
teste <- train(media ~ .,
                        data = df,
                        method = "ranger",
                        importance = "permutation",
                        metric = "Rsquared",
                        trControl = trControl)


teste$resample$Rsquared
mean(teste$resample$Rsquared)
```

```{r}
varImp(teste)
```

```{r}
set.seed(231192)
model_v2 <- train(media ~ nivel_socio_economico + pilar_letramento + tx_resp_q005_F + tx_resp_q017e_A + tx_resp_q010d_B + tx_resp_q017a_A + tx_resp_q004_F + pilar_bens + tx_resp_q004_F + tx_resp_q017e_D + tx_resp_q010i_B + tx_resp_q018a_A + tx_resp_q009c_A + tx_resp_q008b_B,
                        data = df3,
                        method = "ranger",
                        metric = "Rsquared",
                        trControl = trControl)


model_v2$resample$Rsquared
mean(model_v2$resample$Rsquared)
```